Robotic Grasping in Clutter: Using a Hierarchy of Supervisors for Deep Learning from Demonstrations

### 2016-03-16 paper.tex
added figures and updated to draft v3

\title{A Gambler's DAgger: Reducing Supervisor Burden in Online Learning From Demonstration Learning  Through Hierarchical Supervisors}

\author{Michael Laskey$^1$, Jonathan Lee$^1$, Caleb Chuck$^1$, David Gealy$^1$,...,Florian T. Pokorny,\\
 Anca D. Dragan$^1$, and Ken Goldberg$^{1,2}$% <-this % stops a space
\thanks{$^1$ Department of Electrical Engineering and Computer Sciences; {\small \{mdlaskey,iamwesleyhsieh,ftpokorny,anca\}@berkeley.edu, \small staszass@rose-hulman.edu} }%
\thanks{$^2$ Department of Industrial Engineering and Operations Research; {\small goldberg@berkeley.edu}}%
\thanks{$^{1-2}$ University of California, Berkeley;  Berkeley, CA 94720, USA}%
}
\begin{document}

@@

\begin{abstract}
Online learning from demonstration algorithms such as
DAgger can learn policies for problems where the system dynamics and
the cost function are unknown. However they impose a
burden on supervisors to respond to queries each time the robot
encounters new states while executing its current best policy.  Traditionally there has only been one supervisor and it is deemed expert. However, an expert supervisor might not be needed for a large portion of the state space, which could be handled by less-skilled supervisors. We present the Gambler's DAgger, which instead of only using an expert supervisors presents a hierarchical supervisor model, which first leverages cheaper supervisors, such as analytical methods on approximate dynamics or crowdsourced Amazon Mechanical Turk workers to train the policy and then only consults an expert for harder to learn states. We demonstrate this approach in training a visuomotor deep network policy for 2D grasping in clutter on a 4-DOF Zymark Robot. We were able to reduce the number of queries to the final expert supervisor by $X$ and manage to collect $X$K examples by leveraging the hierarchical supervisor model. 
 \end{abstract}

@@

\section{Introduction} 
In model-free robot Learning from Demonstration (LfD), a robot learns to perform a task, such as driving or grasping an object in a cluttered environment, from examples provided by a  supervisor, usually a human.  In such problems, the robot does not have access to either the cost function that it should optimize, nor the dynamics model. The former  when it is difficult to specify how to trade-off various aspects that matter, like a car trying to drive on the road with other cars~\cite{abbeel2004apprenticeship}.  The latter occurs when either the system or the interaction with the world is difficult to characterize, like when a robot is trying to grasp an object in clutter and does not have an accurate model of the object dynamics.

Rather than explicitly learning the cost function as in Inverse Reinforcement Learning and the dynamics model, and then using optimization to produce a policy for the task, in model-free LfD the robot learns a policy directly from supervisor examples, mapping states to controls~\cite{argall2009survey}. Learning from demonstration has been used successfully in recent year for a large number of robotic tasks, including helicopter maneuvering~\cite{abbeel2007application}, car parking~\cite{abbeel2008apprenticeship},  and robot surgery~\cite{van2010superhuman}.

 \begin{figure}
\includegraphics[width=0.5\textwidth]{f_figs/teaser.pdf}
\caption{Typical Initial States for our grasping in clutter task with a Zymark Robot. Red shapes indicate the clutter objects and the yellow circle indicates the goal object. The robot learns from raw image date to push the shapes and reach the yellow circle through our algorithm A Gambler's DAgger. The inscribed circle in the workspace prevents the robot from learning the trivial solution of simply pushing the red objects off the table.}
\vspace*{-20pt}
\label{fig:teaser}
\end{figure}

LfD algorithms can be categorized as offline, where the robot only observes demonstrations, or online, where the robot interacts with the world and receives feedback from the supervisor.
In offline LfD, the robot learns the policy based on a batch of examples, and then executes it to achieve the task.  During execution, a small control error can accumulate, leading the robot away from the region of the state space where it was given examples, leading to unrecoverable failures.  For example, a robot driving may be trained on examples driving safely down the center of a lane, but even slight deviations will eventually put the robot into states near the side of the road where its policy could  fail~\cite{pomerleau1989alvinn}. Ross and Bagnell showed the number of errors made by the robot, in the worst case,  can scale quadratically with the time horizon of the task~\cite{ross2010efficient}.

@@

One drawback is that DAgger significantly increases the burden on the supervisor, who must label all states that the robot visits during training.  Traditionally there has only been one supervisor and it is deemed an expert ~\cite{ross2010efficient,ross2010reduction,ross2013learning,duvallet2013imitation}. However, an expert supervisor might not be needed for a large portion of the state space, which could be handled by less-skilled supervisors.

Our key insight is that for a large portion of the task the supervisor's time is used teaching the robot something that a less costly supervisor could do. For instance in our grasping in clutter example, the robot spends a large portion of the time trying to learn how to go towards the goal object. Analytical techniques such as forward kinematics, template matching and motion planning are readily available to provide examples in these parts of the state space. However, traditional Online Lfd method exploit the expert supervisor for all parts of the state space.

We propose bootstrapping a policy from a hierarchy of supervisors ranked by cost. Examples of such a hierarchy are analytical models with a simplified dynamics of the world, crowd-source Amazon Mechanical Turk workers and then finally an expert supervisor. We demonstrate this approach in training a deep neural network for the visuo-motor based task of 2D grasping in clutter on a Zymark Robot. Our robot consists of an 3DOF arm that lies in a planar workspace and the ability to rotate the table. The objects dataset consists of 20 Medium Density Fiber extruded polygons taken from 2D projections of objects in the Dex-Net \cite{} database.  Results show that by bootstrapping from less skilled experts, the total amount of burden on the expert supervisor is reduced and the task is successfully learned \mlnote{We need an objective measure of success}.

\section{Related Work}
Below we summarize related work in reducing the number of queries to an expert in the Online LfD setting.\mlnote{Still a work in progress}

\noindent \textbf{Online Lfd with an Expert Supervisor}
Successful robotic examples of Online Learning From Demonstration where an expert have been employed have been in flying a quad-copter through a foster, navigating a wheel chair across a room and teaching a robot to follow verbal instructions and surgical needle insertion \cite{ross2013learning, kim2013maximum, duvallet2013imitation, laskey}.

However, to date these approaches have all used only an expert supervisor to label all parts of the state space. Our contribution consists of the use of a hierarchical structure of non-expert supervisors to reduce queries to the expert at the top of the hierarchy.

\noindent\textbf{Reducing Supervisor Burden in Online LfD} One approach that has been studied to reducing supervisor burden is applying active learning to only ask for queries one the robot is uncertain about the correct control to apply. Traditional active learning techniques like query-by-committee and uncertainty sampling have been applied by \cite{chernova2009interactive,judah2011active,grollman2007dogged}

Kim et al. demonstrated that due to the non-stationarity of the problem traditional active learning techniques do not work because the underlying state distribution changes and thus the use of novelty detection is needed to be more conservative in determining uncertainty~\cite{kim2013maximum}.

Laskey et al. created an approach, known as SHIV, for using active learning tailored to high dimensional and non-stationarity state distributions and  provided a modified version One Class SVM to reduce the problem to regularized binary classification~\cite{laskey}. However, the result that novelty detection is needed to reduce supervisor burden via active learning implies that for problems with high stochasticity, such as grasping in clutter with a large number of objects, the expert would be queried a significant number. Furthermore, with the advent of deep learning where the feature space changes each iteration, it becomes unclear as to what feature space to provide novelty detection in.

\noindent\textbf{Training Policies Via Analytical Methods} Instead of relying on a costly human supervisor an alternative approach is to use knowledge of the underlying dynamics and train a policy via a computationally expensive analytical method such as a controller in a physics simulator~\cite{IGOR} or dynamic programming in a Atari video game~\cite{NIPS2014_5421}.

While this approach can "compile" the computationally expensive option into an quickly computed function, it can be limited when analytical models cannot explicitly model the environment. For instance in our grasping in clutter example, the robot has no knowledge of the mass or friction properties of the object, which would make this a challenging problem for such analytical methods. We thus propose using analytical methods not as the expert supervisor, but instead part of a hierarchy where they can be used to reduce the burden of the supervisors above them in the hierarchy.

@@

\noindent\textbf{Modeling Choices and Assumptions}  We model the system dynamics as Markovian, stochastic, and stationary. Stationary dynamics occur when, given a state and a control, the probability of the next state does not change over time. Note this is different from the non-stationary distribution over the states the robot encounters during learning.

We model the initial state as sampled from a distribution over the state space.
We assume a known state space and set of controls. We also assume access to a robot or simulator, such that we  can sample from the state sequences induced by a sequence of controls.   Lastly, we assume access to a set of supervisors who can, given a state, provide a control signal label. We additionally assume the supervisors can be noisy and imperfect, noting that the robot cannot surpass the level of performance of the current supervisor it is training with.

\noindent\textbf{Policies and State Densities.}
Following conventions from control theory, we denote by $\mathcal{X}$ the set consisting of observable states for a robot task, consisting, for example, of 
high-dimensional vectors corresponding to images from a camera, or robot joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of allowable control inputs for the robot, which can be discrete or
continuous. We model dynamics as Markovian, such that the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in
\mathcal{U}$: 
$$p(\bx_{t+1}|\bu_{t},\bx_{t}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{t+1}|\bu_{t}, \bx_t)$$
We assume a probability density over initial states $p(\bx_0)$ that can be sampled from.
%We denote the probability density over the initial state also by $p:\mathcal{X}\to \mathbb{R}$.

A trajectory $\hat{\tau}$ is a finite series of $T+1$ pairs of states visited and corresponding
control inputs at these states, $\hat{\tau} = (\mathbf{x}_0,\mathbf{u}_0, ...., \mathbf{x}_T,\mathbf{u}_T)$, where $\bx_t\in \mathcal{X}$
and $\bu_t\in \mathcal{U}$ for $t\in \{0, \ldots, T\}$ and some $T\in \mathbb{N}$.  
For a given trajectory $\hat{\tau}$ as above, we denote by ${\tau}$ the corresponding trajectory in state space,
${\tau} = (\bx_0,....,\bx_T)$.

@@

While a reward function exists for a task like grasping in clutter, such as a positive value for successfully reaching the goal object. It is often too delayed to provide useful information for quickly learning \cite{kitaevphysics}. We thus assume access to 
a supervisor that can achieve a desired level of performance on the task. The supervisor provides the robot an initial set
 of $N$ stochastic demonstration trajectories $\lbrace \tilde{\tau}^1,...,\tilde{\tau}^N \rbrace$. 
which are the result of the supervisor applying their policy. This induces a training data set $\mathcal{D}$ of all state-control input pairs from the demonstrated trajectories.

@@

Given a candidate policy $\pi_{\theta}$, we then use the surrogate loss function to approximately measure how `close' the policy's
returned control input $\pi_{\theta}(\bx)\in \mathcal{U}$ at a given state $\bx\in \mathcal{X}$ is to the supervisor's policy's control output
$\tilde{\pi}(\bx)\in \mathcal{U}$. The goal is to produce a policy that minimizes the surrogate loss relative to the supervisor's policy.

Following~\cite{ross2010reduction}, our objective is to find a policy $\pi_{\theta}$ minimizing the expected surrogate loss, where the expectation is taken over the distribution of states induced by the policy across any time point in the horizon:

@@

 If the robot could learn the policy  perfectly, this state density would match the one encountered in user examples. But if the robot makes an error, that error changes the distribution of states that the robot will visit, which can lead to states that are far away from any examples and difficult to generalize to~\cite{pomerleau1989alvinn}. This motivates iterative algorithms like DAgger, which iterate between learning a policy and the supervisor providing feedback. The feedback is in the form of control signals on states sampled from the robot's new distribution of states.

 \section{A Gambler's DAgger}
 In this section, we first recall the original DAgger algorithm. Then introduce the notion and formal definition of a hierarchical supervisor and finally describe how to combine DAgger with a hierarchical supervisor to create a Gambler's DAgger.

@@

\subsection{Hierarchy of Supervisors}
Instead of one expert supervisor,we propose a hierarchy  of $M$ supervisors where for each supervisor $\tilde{\pi}_m$, there is an associated expected cumulative reward $R_m = \int \sum^T_{t=1} r(\mathbf{x}_t,\mathbf{u}_t) p(\mathbf{x}|\tilde{\pi}_m)d\mathbf{x})$, where $p(\mathbf{x}|\tilde{\pi}_m)$ denotes the average distribution of states the supervisor encounters. We also denote some measure of cost $C_m$ that is ascribed to the supervisor, such as computational time or monetary expense. We assume that the rank of supervisors in terms of cost is equivalent to the rank of supervisor in the terms of expected cumulative reward. Thus, the cheapest supervisor also receives the lowest expected cumulative reward.

Additionally for two supervisors $\tilde{\pi}_m$ and $\tilde{\pi}_{m+1}$ to be in the hierarchy, they must only disagree in expectation on a subset of the work space to some precision $\tilde{\epsilon}$. Denote the set of states two supervisors disagree as $\mathcal{X}_{m,m+1} = \lbrace \mathbf{x} | ||\tilde{\pi}_m(\mathbf{x}) - \tilde{\pi}_{m+1}(\mathbf{x}) ||^2_2 > \tilde{\epsilon} \rbrace$ .We make this condition because if two supervisors never provided the same examples then all states would have to be relabeled by the next supervisor in the hierarchy.

We formally define two supervisor's in a hierarchy as follows

\begin{definition} Two supervisors $\tilde{\pi}_m$ and $\tilde{\pi}_{m+1}$ are in a hierarchy if  $C_m < C_{m+1}$ , $R_m < R_{m+1}$ and  $\mathcal{X}_{m,m+1} \subset \mathcal{X}$
\end{definition}

We note that a two hierarchical supervisors according to this definition also implies that supervisor $\tilde{\pi}_{m+1}$ in the hierarchy is higher than the set of supervisors $\lbrace \tilde{\pi}_i | \forall i \in [0,m] \rbrace$.

\subsection{A Gambler's DAgger}
Our algorithm is as follows: first iterate through Step 1 and 2 of DAgger for a given supervisor, however after iteration $K$ or when the current policy, $\pi_{\theta_{m,k}}$ is able to achieve a loss $\epsilon$ on the sampled $J$ states at that current iteration (i.e):

@@

Then iterate to the next supervisor in the hierarchy or $m = m+1$. An issue arises though when performing this iterations because now the current dataset $\mathcal{D}_K$ has examples from a different supervisor that receives a smaller expected cumulative reward. This could cause a learning algorithm to try and regress to both labels and diverge from the intended performance \cite{scholkopf2002learning}.

We are thus interested in not storing the dataset collected from the previous supervisor and only remembering the current $\theta_{m,K}$ weight parameters. Thus, we propose only performing DAgger for each individual supervisor in the hierarchy and using the resulting weight vector to bootstrap the learning of the next supervisor. \mlnote{the last step about the dataset is something I want to empirically validate}

>>>>

\title{Robotic Grasping in Clutter: Using a Hierarchy of Supervisors for Deep Learning from Demonstrations}

\author{Michael Laskey$^1$, Jonathan Lee$^1$, Caleb Chuck$^1$, David Gealy$^1$,Wesley Hseih$^1$,Florian T. Pokorny,\\
 Anca D. Dragan$^1$, and Ken Goldberg$^{1,2}$% <-this % stops a space
\thanks{$^1$ Department of Electrical Engineering and Computer Sciences; {\small \{mdlaskey,iamwesleyhsieh,ftpokorny,anca\}@berkeley.edu, \small staszass@rose-hulman.edu} }%
\thanks{$^2$ Department of Industrial Engineering and Operations Research; {\small goldberg@berkeley.edu}}%
\thanks{$^{1-2}$ University of California, Berkeley;  Berkeley, CA 94720, USA}%
}
\begin{document}

@@

\begin{abstract}
Online learning from demonstration algorithms such as
DAgger can learn policies for problems where the system dynamics and
the cost function are unknown. However they impose a
burden on supervisors to respond to queries each time the robot
encounters new states while executing its current best policy.  Traditionally there has only been one supervisor and it is an expert. However, an expert supervisor is a costly resource that could be impractical for large scale learning tasks. However, there an expert supervisor might not be needed for a large portion of the state space , which could be handled by less-skilled supervisors. We present the Gambler's DAgger, which instead of only using an expert supervisors presents a hierarchical supervisor model, that first leverages cheaper supervisors, such as analytical methods on approximate dynamics or crowdsourced Amazon Mechanical Turk workers to train the policy. Then only consults an expert supervisor for harder to learn states. We demonstrate this approach in training a visuomotor deep network policy in TensorFlow for grasping in clutter on a 4-DOF Zymark Robot. We were able to reduce the number of queries to the final expert supervisor by $X$ and manage to collect $X$K examples by leveraging the hierarchical supervisor model. 
 \end{abstract}

@@

\section{Introduction} 
Consider a robot trying to reach an object on a cluttered shelf in a shipping warehouse, as illustrated by the Amazon Picking Challenge at ICRA 2015 \cite{mahler2016dexnet}, this is an important task for being able to efficiently ship and process orders.   The robot is presented with a visual feed of the environment and must plan a series of motions to push the cluttered obstacles and grasp a goal object. A challenge could arise  when the robot encounters an unknown obstacle object that it does not have a known dynamics  for and must manipulate the environment to reach goal.

One solution is to learn a manipulation policy across a variety of obstacle objects and try to generalize this approach to unknown obstacle in the environment. Reinforcement learning techniques are capable of being applied to this problem, however they could require a large amount of trials on the robot to accomplish the task when the dynamics are unknown and the state space is high dimensional image data~\cite{pinto2015supersizing}.

Rather than using an unsupervised learning approach, we propose guiding the robot how on to interact with the environment via a supervisor, who knows how to accomplish the task~\cite{argall2009survey}. Learning from demonstration (LfD) algorithms, which have been used successfully in recent year for a large number of robotic tasks, including helicopter maneuvering~\cite{abbeel2007application}, car parking~\cite{abbeel2008apprenticeship},  and robot surgery~\cite{van2010superhuman}, is a form of how to provide this guidance.

 \begin{figure}
\includegraphics[width=0.5\textwidth]{f_figs/teaser.eps}
\caption{Typical Initial States for our grasping in clutter task with a Zymark Robot. Red shapes indicate the clutter objects and the yellow circle indicates the goal object. The robot learns from 250x250 RGB image data to push the shapes and reach the yellow circle through our algorithm A Gambler's DAgger.}
\vspace*{-20pt}
\label{fig:teaser}
\end{figure}

LfD algorithms can be categorized as offline, where the robot only observes demonstrations, or online, where the robot interacts with the world and receives feedback from the supervisor.
In offline LfD, the robot learns the policy based on a batch of examples, and then executes it to achieve the task.  During execution, a small control error can accumulate, leading the robot away from the region of the state space where it was given examples, leading to unrecoverable failures.  For example, a robot  trained on examples driving safely down the center of a lane, but even slight deviations will eventually put the robot into states near the side of the road where its policy could  fail~\cite{pomerleau1989alvinn}. Ross and Bagnell showed the number of errors made by the robot, in the worst case,  can scale quadratically with the time horizon of the task~\cite{ross2010efficient}.

@@

One drawback is that DAgger imposes a substantial burden on the supervisor, who must label all states that the robot visits during training.  Traditionally there has only been one supervisor and it is deemed an expert ~\cite{ross2010efficient,ross2010reduction,ross2013learning,duvallet2013imitation}. However, when learning complicated tasks with high stochasticity in the environment this could require a large amount of the expert's time, which is a costly resource.

Learning can be bootstrapped with a sequence of increasingly skilled supervisors. In our grasping in clutter example, the robot spends a large portion of the time learning how to go towards the goal object. Analytical techniques such as forward kinematics, template matching and motion planning are readily available to provide examples in these parts of the state space.

We propose bootstrapping a policy from a hierarchy of supervisors ranked by cost: analytical models, crowd-sourced workers and an expert supervisor to learn the robot grasping in clutter task. We demonstrate this approach by training a deep neural network for the visuo-motor based task of  grasping in clutter on a Zymark Robot. Our robot consists of an 2DOF arm and a gripper that lie in a planar workspace. The objects dataset consists of 20  extruded polygons taken from 2D projections of objects in the Dex-Net \cite{mahler2016dexnet} database.  Results show that by leveraging a hierarchy of less skilled experts, the total amount of cost incurred is reduced by X$\%$ and the task is $61\%$ successful on our test set of 20 object configurations of shapes not trained on \mlnote{need to think of a better teaser statement}

\section{Related Work}
Below we summarize related work in Robotic Grasiping in Clutter,  Online LfD setting and then the field of Curriculum learning, which is similar to the concept of hierarchical supervisors.

\noindent \textbf{Robotic Grasping in Clutter}
Robotic grasping has been extensively studied over the
past four decades~\cite{bicchi2000robotic}. A significant amount of prior work focuses on planning grasps given a known object.  However, in unstructured environments clutter poses a significant challenge for reaching the planned grasp ~\cite{katz2008can}.
Prior work has addressed integrated perception and grasping
in clutter where the objective is to grasp objects in an
unorganized pile ~\cite{pinto2015supersizing,nieuwenhuisen2013mobile}, but these
methods do not specifically aim to grasp a single target object
in clutter. 
Leeper et al. ~\cite{} use a human operator for assistance to guide the
robot through clutter with the objective of grasping a target
object. However, the robot required the human to be present at all times and did not attempt to learn to operate autonomously.We are interested in only querying the supervisor for examples until the robot is successful at the task.

Prior work has studied the issue of manipulating objects
by performing pushing operations~\cite{}. Berenson et
al. ~\cite{} use a sampling-based planner that considers clearance
from obstacles in the environment to plan a grasp
approach trajectory in cluttered environments. Cosgun et al.
[3] and King et al. \cite{kingnonprehensile} consider the problem of planning
a series of push operations that would place an object at a
desired target location. Kiteav et al. planned a trajectory in a physics simulator using LQG based controllers \cite{kitaevphysics}. However, all of these works assume a known model of the world is given to them, we are interested in learning manipulation polices via leveraging an supervisors intuition for the the task.  \mlnote{We should discuss how to distinguish our work (Properties of our approach are we assume raw image data, real time and unknown dynamics)}.

\noindent \textbf{Online Lfd with an Expert Supervisor}
Successful robotic examples of Online Learning From Demonstration where an expert has been employed in flying a quad-copter through a forest, navigating a wheel chair across a room and teaching a robot to follow verbal instructions and surgical needle insertion \cite{ross2013learning, kim2013maximum, duvallet2013imitation, laskey}.

However, to date these approaches have all used only one expert supervisor to provide examples for all parts of the state space. Our contribution consists of the use of a hierarchical structure of different skill level supervisors to reduce learning cost.

\noindent\textbf{Reducing Supervisor Burden in Online LfD} One approach that has been studied to reducing queries to the supervisor is applying active learning to only ask for queries one the robot is uncertain about the correct control to apply. Traditional active learning techniques like query-by-committee and uncertainty sampling have been applied by \cite{chernova2009interactive,judah2011active,grollman2007dogged}

However, Kim et al. demonstrated that due to the non-stationarity of the problem traditional active learning techniques may not work because the underlying state distribution changes and thus the use of novelty detection is needed to be more conservative in determining uncertainty~\cite{kim2013maximum}. Laskey et al. introduced SHIV, for using active learning tailored to high dimensional and non-stationarity state distributions and  provided a modified version One Class SVM to reduce the from density estimation, which requires regression, to the simpler regularized binary classification ~\cite{laskey}.

Novelty detection being needed to reduce supervisor burden via active learning implies that for problems with high stochasticity, such as grasping in clutter, the expert would be queried a significant number. Furthermore, we are interested in using deep neural architectures to learn task specific features for the task. SHIV applies novelty detection in a known feature space, however the feature space now changes each iteration. Thus, it is not straightforward to apply existing techniques.

\noindent\textbf{Curriculum Learning}
\mlnote{TODO: add more curriculum}
Our approach is similar to the area known as curriculum learning, where a neural network is trained via first  easier to learn examples and then gradually increasing the difficulty of the examples to learn ~\cite{bengio2009curriculum}.

Sanger et al. used curriculum learning in robotics to gradually train a neural network policy to learn the inverse dynamics of a robot manipulator. They then varied then data collection scheme such that easily learned trajectories where shown first and then gradually increased the difficulty~\cite{sanger1994neural}.

Our approach is different than curriculum learning in that a supervisor can be provide low quality examples in parts of the state space visited by the robot, which means it isn't neccesarily easier for the robot to learn one supervisor's policy versus another in the hierarchy. Combining hierarchical supervisors and curriculum learning could be a promising avenue for future work though.

@@

\noindent\textbf{Assumptions and Modeling Choices } 
We assume a known state space and set of controls. We also assume access to a robot or simulator, such that we  can sample from the state sequences induced by a sequence of controls.Lastly, we assume access to a set of supervisors who can, given a state, provide a control signal label. We additionally assume the supervisors can be noisy and imperfect, noting that a lower cost supervisor also has lower quality.

 We model the system dynamics as Markovian, stochastic, and stationary. Stationary dynamics occur when, given a state and a control, the probability of the next state does not change over time. Note this is different from the non-stationary distribution over the states the robot encounters during learning. We model the initial state as sampled from a distribution over the state space.

\noindent\textbf{Policies and State Densities.}
Following conventions from control theory, we denote by $\mathcal{X}$ the set consisting of observable states for a robot task, consisting, for example, of 
high-dimensional vectors corresponding to images from a camera, or robot joint angles and object poses in the environment.
We denote by $\mathcal{U}$ a set of allowable control inputs for the robot, which can be discrete or
continuous. We model dynamics as Markovian: the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in
\mathcal{U}$: 
$$p(\bx_{t+1}|\bu_{t},\bx_{t}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{t+1}|\bu_{t}, \bx_t)$$
We assume an unknown probability density over initial states $p(\bx_0)$.
%We denote the probability density over the initial state also by $p:\mathcal{X}\to \mathbb{R}$.

A demonstration (or trajectory )$\hat{\tau}$ is a series of $T+1$ pairs of states visited and corresponding
control inputs at these states, $\hat{\tau} = (\mathbf{x}_0,\mathbf{u}_0, ...., \mathbf{x}_T,\mathbf{u}_T)$, where $\bx_t\in \mathcal{X}$
and $\bu_t\in \mathcal{U}$ for $t\in \{0, \ldots, T\}$ and some $T\in \mathbb{N}$.  
For a given trajectory $\hat{\tau}$ as above, we denote by ${\tau}$ the corresponding trajectory in state space,
${\tau} = (\bx_0,....,\bx_T)$.

@@

The reward function  for a task like grasping in clutter, (i.e.  successfully grasping the goal object) is delayed \cite{kitaevphysics}. The idea behind DAgger and SHIV is to query a supervisor for appropiate actions, to provide the robot a set of N stochastic demonstrations trajectories $\lbrace \hat{\tau}^1,...\hat{\tau}^N \rbrace$. This induces a training data set $\mathcal{D}$ of state-control input pairs.

@@

Given a candidate policy $\pi_{\theta}$, we then use the surrogate loss function to approximately measure how `close' the robot's policy's
returned control input $\pi_{\theta}(\bx)\in \mathcal{U}$ at a given state $\bx\in \mathcal{X}$ is to the supervisor's policy's control output
$\tilde{\pi}(\bx)\in \mathcal{U}$. The goal is to produce a policy that minimizes the surrogate loss.

Following~\cite{ross2010reduction}, our objective is to find a policy $\pi_{\theta}$ minimizing the expected surrogate loss with respect to the most skilled supervisor, where the expectation is taken over the distribution of states induced by the policy across any time point in the horizon:

@@

 If the robot could learn the policy  perfectly, this state density would match the one encountered in user examples. But if the robot makes an error, that error changes the distribution of states that the robot will visit, which can lead to states that are far away from any examples and difficult to generalize to~\cite{pomerleau1989alvinn}. This motivates iterative algorithms like DAgger, which iterate between learning a policy and the supervisor providing examples.

We  introduce the concept of their being a set of supervisors each with an associated cost and skill level. Denote $\tilde{\pi}_M$ the most skilled and costly supervisor in this set. This supervisor could for example be a person with a Phd in machine learning and robotics. We formally define how each supervisor in the set is related to each other in Sec. \ref{sec:hS}.  Every time a supervisor is queried a cost, $C_m$, is incurred this could be an hourly pay rate or computational time.

While Eq. \ref{eq:LFD_obj} is the primary objective of our algorithm a secondary objective is to reduce the cumulative cost associated with training the robot. Thus, we are interested in solving the following objective

 \vspace{-2ex}
\begin{align}\label{eq:bugjet_LFD_obj}
\underset{\theta,C_{m,j}}{\min} \: & \sum^J_{j=1} C_{m,j}\\
 \mbox{s.t.} \: &E_{p(\bx|\theta)} [l(\pi_\theta(\bx),\tilde{\pi}_M(\bx))] \leq \epsilon
\end{align}

where $J$ is the number of queries need to satisfy the constraint on expected surrogate loss, $C_{m,j}$ denotes the cost incurred by querying the supervisor at that iteration and $\epsilon$ is a user-defined threshold on how close robot policy is to match the highest quality supervisor. We present our algorithm, LEATHERMAN, which aims to solve Eq. \ref{eq:bugjet_LFD_obj}.

 \section{Algorithm}
 In this section, we first recall the original DAgger algorithm. Then introduce the notion and formal definition of a hierarchical supervisor and finally describe how to combine DAgger with a hierarchical supervisor to create a Gambler's DAgger.

@@

\subsection{Hierarchy of Supervisors}\label{sec:hS}
Instead of one  supervisor,we propose a hierarchy  of $M$ supervisors where for each supervisor $\tilde{\pi}_m$, there is an associated expected cumulative reward $R_m = \int \sum^T_{t=1} r(\mathbf{x}_t, \tilde{\pi}_m(x)) p(\tau |\tilde{\pi}_m)d\tau$, where $p(\tau |\tilde{\pi}_m)$ denotes the average distribution of states the supervisor encounters. We also denote some measure of cost $C_m$ that is ascribed to the supervisor, such as computational time or monetary expense. Generally the rank of supervisors in terms of cost is equivalent to the rank of supervisor in the terms of expected cumulative reward. Thus, the least costly supervisor will receive the lowest expected cumulative reward.

Additionally for two supervisors $\tilde{\pi}_m$ and $\tilde{\pi}_{m+1}$ to be in the hierarchy, they must only disagree in expectation on a subset of the work space to some precision $\tilde{\epsilon}$. Denote the set of states two supervisors disagree as $\mathcal{X}_{m,m+1} = \lbrace \mathbf{x} | ||\tilde{\pi}_m(\mathbf{x}) - \tilde{\pi}_{m+1}(\mathbf{x}) ||^2_2 > \tilde{\epsilon} \rbrace$ .We enforce the condition $\mathcal{X}_{m,m+1} \subset \mathcal{X}$ because if two supervisors never provided the same examples then all states would have to be relabeled by the next supervisor in the hierarchy.

We formally define an ordering on supervisors:

\begin{definition} Two supervisors $\tilde{\pi}_m$ and $\tilde{\pi}_{m+1}$ are in a hierarchy if  $C_m < C_{m+1}$  and  $\mathcal{X}_{m,m+1} \subset \mathcal{X}$
\end{definition}

Violation of this definition could result in no additional benefit from the LEATHERMAN algorithm or potentially worst a higher budget incurred than without using a hierarchy of supervisors.

\subsection{DAgger with Supervisor Hierarchy}
\mlnote{Anca and I wanted to justify the greedy allocation section, which we draw from the contextual arm bandit literature to do so... I can condense this}
The hierarchical supervisor problem can be framed as a contextual multi-armed bandit, which is like a traditional multi-armed bandit except the sampled reward depends on an additional state term~\cite{li2010contextual}. At each iteration a supervisor is selected from $M$ supervisors  and the state is the current policy parameters or $\theta_k$. Then a sampled reward is received which is how well the trained policy performed on the sample surrogate loss measured against the $\tilde{\pi}_M$ or the supervisor at the top of the hierarchy. Note we only measure against the supervisor at the top of the hierarchy because that is our primary objective (i.e. Eq. \ref{eq:LFD_obj}).

While a large amount of literature exists to solve contextual multi-armed bandit problems, a common assumption is that the sampled reward can be observed~~\cite{li2010contextual}. However, in our situation evaluation of the sampled reward requires querying the costliest supervisor every iteration. Thus, defeating the purpose of cost reduction. In light of this we propose a greedy allocation strategy where we train a policy with the cheapest supervisor first until convergence and then work up the hierarchy training with each supervisor.

Our algorithm is as follows: first iterate through Step 1 and 2 of DAgger for a given supervisor, however after iteration $K$ or when the current policy, $\pi_{\theta_{m,k}}$ is able to achieve a loss $\epsilon$ on the sampled $J$ states at that current iteration (i.e):

@@

Then iterate to the next supervisor in the hierarchy or $m = m+1$. The hyperparameter $\epsilon$ can be set as a measure of how different the controls applied between the robot's policy and the current supervisors a sequential process can tolerate. For example surgical sequential processes could warrant a very low threshold, but less precise sequential processes, like grasping in clutter, can tolerate a higher value.

An issue arises though when performing this iterations because now the current dataset $\mathcal{D}_K$ has examples from a different supervisor that receives a smaller expected cumulative reward. This could cause a learning algorithm to try and fit to contradictory examples  and be worse than either supervisor trained with \cite{scholkopf2002learning}.

We are thus interested in not storing the dataset collected from the previous supervisor and only remembering the current $\theta_{m,K}$ weight parameters. Thus, we propose only performing DAgger for each individual supervisor in the hierarchy and using the resulting weight vector to bootstrap the learning of the next supervisor. We found empirically in Sec. \ref{sec:Exp} that it also beneficial to add to the new dataset the states and controls applied by the current policy $\pi_{\theta_{m,k}}$ that the current supervisor agreed with by some euclidean distance $\tilde{\epsilon}$.Thus, acting as a regularization on the optimization since stochastic gradient on a section of the min-batch update would be zero for examples where the next supervisor agrees with the previous.

\mlnote{we can formally prove the following statements or point to the web similar to SHIV... it will be alot easier to write these out though if we have space}
With respect to the theoretical analysis of DAgger's convergence rate, one can see that in our algorithm you are solving $M$ individual DAgger's ~\cite{ross2010reduction}. Thus, in the worst case you will converge to the supervisor at the top of the hierarchy $\tilde{\pi}_M$ in $\tilde{O}(MT)$ iterations instead of DAgger's original $\tilde{O}(T)$.

It should be noted though that as you iterate through the hierarchy of supervisors, each iteration starts out with an error bounded by $\int_{\mathcal{X}_{m,m+1}} ||\pi_{\theta_{0,m+1}}(\mathbf{x}) - \tilde{\pi}_{m+1}(\mathbf{x})||^2_2 p(\mathbf{x}|\theta_{0,m+1}) d\mathbf{x} + \epsilon$. Thus, if the supervisors are able to agree on a large part of the state space, there should be less iterations with the expert supervisor than $\tilde{O}(T)$, which we experimentally show in our grasping in clutter task.

\begin{figure}[t]
\centering

\includegraphics[width=\columnwidth, height=4cm]{f_figs/labeling.pdf}

\caption{\footnotesize  The interface AMT workers see for providing feedback to the robot for the grasping in clutter task. The pink overlay indicates the desire change in the robot position with respect to the current robot state. Then AMT workers can use their intuition for how objects respond to force to provide examples of how the robot should behave. Each window shows a correction in a degree of freedom a) is the rotation of the robots base b) is extending the robot's arm.}
\vspace*{-10pt}
\label{fig:overlays}
\end{figure}

\section{Robot Grasping in Clutter}
We now describe the grasping in clutter problem, the supervisor hierarchy that we used to solve it and our deep learning policy architecture.
